---
title: "Exploring, Assessing PSRC Data"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
  # html_notebook:
  #   toc: yes
  #   toc_float: yes
editor_options:
  chunk_output_type: inline
---

This document will be checking out the PSRC data.


# Initialization

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, root.dir = here::here())
```

```{r, Imports, include=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(janitor)
```

## Code for building tables, etc.

#### Summary table:
```{r, sumtable}
sumtbl <- function(.data, ...) {
  .data %>% 
    group_by(...) %>% 
    summarize(n = n())
}
```

#### Crosstab:
```{r, crosstab}
sum_by_year <- function(data, sumvar) {
  data %>% 
    tabyl({{ sumvar }}, survey_year) %>% 
    adorn_totals(where = c("row", "col")) %>%
    adorn_percentages(denominator = "col") %>% 
      adorn_pct_formatting() %>% 
    adorn_ns()
}

```


# Household Level {.tabset .tabset-fade .tabset-pills}

Import household data:
```{r, message=FALSE}
hhdat <- read_rds(here("analysis/data/derived_data/hhdat.rds"))
```

## other stuff
```{r}
# hhdat %>% 
#   tabyl(final_home_rgcnum, final_home_uvnum)
```


## Basic counts

#### How many households from 2017 vs 2019

```{r, echo = TRUE}
hhdat %>% sumtbl(survey_year)
```

#### Number of smartphone vs travel diary

```{r}
hhdat %>% 
  sumtbl(hhgroup)
```

#### Cross of those two: how many in each survey year

```{r}
hhdat %>% 
  sum_by_year(hhgroup)
```

## Income

#### Broad income (everyone answered)

```{r}
hhdat %>% 
  sum_by_year(hhincome_broad)
```


#### Detailed income

```{r}
hhdat %>% 
  sum_by_year(hhincome_detailed)
```

#### Follow-up categories if they said "prefer not to answer" to the detailed question

```{r}
hhdat %>% 
  filter(!is.na(hhincome_followup)) %>% 
  sum_by_year(hhincome_followup)
```

## Household size

```{r}
hhdat %>% 
  sum_by_year(hhsize)

```

#### Number adults and children

```{r}
# hhdat %>% 
#   mutate(hhsize = as.numeric(str_sub(hhsize, start = 1, end = 2))) %>% 
#   mutate(adult_ratio = numadults/hhsize) %>%  
#   sum_by_year(adult_ratio)

hhdat %>% sum_by_year(numadults)

```

#### Number of workers

```{r}
hhdat %>% 
  sum_by_year(numworkers)
```

#### Number of workers by number of adults for 2017
```{r}
hhdat %>% 
  filter(survey_year == 2017) %>% 
  tabyl(numworkers, numadults) %>% 
  adorn_totals() %>% 
  adorn_percentages(denominator = "col") %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  adorn_title()
```

## Life Cycle

```{r}
hhdat %>% 
  sum_by_year(lifecycle)

```

Is `lifecycle` exclusionary? How could that work? If not, how is it decided which group they are in? 

```{r}
hhdat %>% 
  filter(lifecycle == "Household includes children under 5" | lifecycle == "Household includes children age 5-17") %>% 
  select(hhid, lifecycle, numchildren, numadults) %>% 
  View(title = "kids")

hhdat %>% 
  filter(numchildren == 0) %>% 
  select(hhid, lifecycle, numchildren, numadults) %>% 
  View(title = "nokids")
```

**If they have kids, then they are put into a "kids" group.** The group is determined by the youngest kid they have. If under 5, then they are put in "children under 5" group. If over 5, they are put in "children age 5-17" group. 

Else, they are put into the appropriate group based on number of hh members and householder age. 

## Day of week of diary

```{r}
hhdat %>% 
  sum_by_year(dayofweek)
```
Travel diary **ONLY** Tuesday, Wednesday, Thursday. For "typical day" type of measurement.


## Residence type

#### Rent/own
```{r}
hhdat %>% 
  sum_by_year(rent_own)
```

#### Residence type

```{r}
hhdat %>% 
  sum_by_year(res_type)
```



## Number of weekdays 

Doing this one because I don't understand what it's purpose is.

```{r}
hhdat %>% 
  select(lifecycle, hhsize, nwkdays, dayofweek) %>% 
  head()
```

Still don't understand what this means.


# Person Level {.tabset .tabset-fade .tabset-pills}
What do we have in the person-level data?

Import Person Data
```{r}
prdat <- read_rds(here("analysis/data/derived_data/prdat.rds"))
```

## Other stuff
```{r}
prdat %>% 
  tabyl(student, worker) %>% 
  adorn_percentages() %>% adorn_pct_formatting()
```

```{r}
prdat %>% 
  tabyl(commute_mode, student)
```



## Demographics

#### Age

```{r}
prdat %>% 
  sum_by_year(age)
```

```{r}
prdat %>% 
  sum_by_year(age_category)
```


#### Gender

```{r}
prdat %>% 
  sum_by_year(gender)
```

## SES
Socioeconomic status variables.

### Employment

#### General employment variable
```{r}
prdat %>% 
  sum_by_year(employment)
```

Who are in the `NA` group? I think it's probably ppl under 18
```{r}
prdat %>% 
  filter(is.na(employment)) %>%
  pull(age_category) %>% 
  unique()

prdat %>% 
  filter(age_category == "Under 18 years") %>% 
  pull(employment) %>%
  unique()
```

Are number under 18 and number `NA` for employment the same?
```{r}
prdat %>% filter(is.na(employment)) %>% pull(employment) %>% length()
prdat %>% filter(age_category == "Under 18 years") %>% pull(age_category) %>% length()
```
YES. All `NA` values for `employment` are ppl under 18. 


#### Number of jobs

```{r}
prdat %>% 
  sum_by_year(jobs_count)
```


#### Worker

```{r}
prdat %>% 
  sum_by_year(worker)
```

```{r}
prdat %>% 
  filter(age == "16-17 years" | age_category != "Under 18 years") %>% 
  filter(worker == "No jobs") %>%
  select(hhid, person_id, age, employment, jobs_count, worker) %>% 
  View(title = "non-workers")
```


```{r}
prdat %>% 
  tabyl(jobs_count, worker)
```

```{r}
prdat %>% 
  tabyl(employment, worker)
```


Difference between `worker` and `jobs_count`: `jobs_count` is only "0 jobs" for 16-17 year-olds. Otherwise, it is `NA`. The values for both `jobs_count` and `worker` are determined by the `employment` variable. If `employment` is "retired", "homemaker", or "not currently employed", then `jobs_count` is `NA`, and  `worker` is "No jobs". 

For my analysis, `worker` is probably useful enough on it's own. However, `jobs_count` is interesting. If ppl have multiple jobs in the same day, they need to go to more places. This could pose issues with taking public transit or other modes besides SOV's, where freedom of movement/flexibility is reduced.


## Education
#### Education

```{r}
prdat %>% 
  sum_by_year(education)
```

```{r}
prdat %>% 
  tabyl(education, age_category) %>% 
  adorn_percentages(denominator = "col") %>% adorn_pct_formatting()
```
Seattle area has skewed demographics in terms of educational attainment: number of people with bachelor degree and above is extremely high. In particular, number with graduate/post-graduate degree.

```{r}
prdat %>% 
  sum_by_year(education)
```



#### Student
```{r}
prdat %>% 
  sum_by_year(student)
  
```

```{r}
prdat %>% 
  tabyl(student, age_category)
```

School type

```{r}
prdat %>% 
  sum_by_year(schooltype)
```

```{r}
prdat %>% 
  tabyl(schooltype, age_category) %>% 
  adorn_percentages(denominator = "col") %>% adorn_pct_formatting()
```

## Travel Behavior
#### Workplace

```{r}
prdat %>% 
  sum_by_year(workplace)
```
#### License
```{r}
prdat %>% 
  sum_by_year(license)
```
#### Commute

Frequency:
```{r}
prdat %>% 
  sum_by_year(commute_freq)
```

```{r}
prdat %>% 
  sum_by_year(commute_mode)
```

```{r}
prdat %>% 
  sum_by_year(telecommute_freq)
```

```{r}
prdat %>% 
  sum_by_year(school_freq)
```
School frequency was NOT repeated in 2019.

#### Mode

Times ridden transit in past 30 days
```{r}
prdat %>% 
  sum_by_year(mode_freq_1)
```

Times ridden a bike in past 30 days
```{r}
prdat %>% 
  sum_by_year(mode_freq_2)
```

Times gone for a walk in past 30 days

```{r}
prdat %>% 
  sum_by_year(mode_freq_3)
```

Times used carshare in past 30 days
```{r}
prdat %>% 
  sum_by_year(mode_freq_4)
```


Times used rideshare in past 30 days
```{r}
prdat %>% 
  sum_by_year(mode_freq_5)
```


## Attitudes, opinions

### Preference survey
```{r}
prdat %>% 
  sum_by_year(completed_pref_survey)
```
```{r}
prdat %>% 
  filter(!(completed_pref_survey == "Yes")) %>% 
  View()
```

This was only included in 2017. 

According to p. 19 of the PSRC 2017 final report, preference survey was the following:

* various mode frequencies (how often one typically uses modes)
* (If uses transit) availability and use of varioustransit fare payment methods
* Employer transit subsidies and commuter benefits
* Autonomous vehicle concerns and interests
* Factors that would encourage increased bike/transit usage

If a person's survey was reported by proxy, then they were not asked opinion and preference questions. This consisted of `211` respondents in 2017.

#### Use more transit

Safer ways to get to stops
```{r}
prdat %>% 
  sum_by_year(wbt_transitmore_1)
```

Increased frequency
```{r}
prdat %>% 
  sum_by_year(wbt_transitmore_2)
```


Increased reliability
```{r}
prdat %>% 
  sum_by_year(wbt_transitmore_3)
```



#### Use more bike
Shared use path or protected bike lane
```{r}
prdat %>% 
  sum_by_year(wbt_bikemore_1)
```

Neighborhood greenway
```{r}
prdat %>% 
  sum_by_year(wbt_bikemore_2)
```

Bike lane
```{r}
prdat %>% 
  sum_by_year(wbt_bikemore_3)
```


Shared roadway lane
```{r}
prdat %>% 
  sum_by_year(wbt_bikemore_4)
```

End of trip amenities
```{r}
prdat %>% 
  sum_by_year(wbt_bikemore_5)
```

# Trip Level {.tabset .tabset-fade .tabset-pills}

Import Trip-level Data
```{r}
trdat <- read_rds(here("analysis/data/derived_data/trdat.rds"))
```



## Diary day sequence

Not sure what this is so i'm checking it out

```{r}
# trdat %>% pull(personid)
trdat %>% 
  filter(personid == 1710065203) %>% 
  pull(daynum) %>% 
  unique()
```
SO: this person (id: `1710065203`) has 7 daynums recorded. According to the psrc 2017 final report: those that used rSurvey reported their travel for 1 day. Those that used rMove reported their travel for 7 days. 

It also says that households were "preassigned" to a Tues, Wed, or Thurs, and then reassigned to week always starting with a Tues if they were using rMove. 


## Mode stuff

```{r}
modedat <- trdat %>% 
  select(hhid, personid, contains("mode"))
```


```{r}
walk2bus <- trdat %>% 
  filter(str_detect(mode_1, "Walk")) %>% 
  filter(str_detect(mode_2, regex("bus", ignore_case = TRUE)) | 
           str_detect(mode_2, regex("rail", ignore_case = TRUE)) 
         )
```

Which trips even have multiple modes?
```{r}
multmode <- modedat %>% 
  filter(!is.na(mode_2))
```

1,584 trips out of 124,516 have multiple modes. Out of those, 


```{r}
multmode %>% 
  group_by(mode_1) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```

```{r}
multmode %>% 
  tabyl(mode_1) %>% 
  arrange(desc(n))
```

```{r}
multmode %>% 
  tabyl(mode_2) %>% 
  arrange(desc(n)) %>% 
  adorn_pct_formatting()
```

```{r}
trdat %>% 
  tabyl(mode_1) %>% 
  arrange(desc(n)) %>% 
  adorn_pct_formatting()
```


```{r}
multmode %>% 
  group_by(mode_2) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```



## First travel beh LPA: Main mode

```{r}
trdat %>% 
  group_by(main_mode) %>% 
  summarise(n = n())
```



Get all the modes used by each person . Should look like this:


| personid    | Bike    | Walk    | Transit   | HOV     | SOV     | Other     |
|-------------|---------|---------|-----------|---------|---------|-----------|
| 12345       | 1       | 1       | 1         | 0       | 1       | 1         |
| 23456       | 0       | 1       | 1         | 0       | 1       | 0         |


This is done in `analysis/04_prep-mode-data.r`





